### 2020-07-11 paper/sex_selection_spacing-ver4.tex
Clarify transformation of data and sex ratio calculation
- use "interval" rather than period
- better wording for transformation
- reorganize description of sex ratio calculation method

Equation (\ref{eq:hazard}) is equivalent to the logistic hazard model and has the same 
likelihood function as the multinomial logit model \citep{allison82,jenkins95}.
Hence, transforming the data, so each observation is a time period---here equal
to three months---I can estimate the model using a standard multinomial logit model.
%
% \footnote{
% The multinomial model assumes that alternative exit states are stochastically independent
% (Independence of Irrelevant Alternatives or IIA), which rules out any individual-specific 
% unmeasured or unobservable factors that affect both the hazard of having a girl and the 
% hazard of having a boy.
% I, therefore, include a proxy for fecundity discussed in Section \ref{sec:data}.
% Also, the multivariate probit model can be used as an alternative because the IIA is not 
% imposed \citep{han90}.
% The results are substantially identical between these two models and
% available upon request.
% }
In the reorganized data the outcome variable is 0 if the woman does not have a child in a 
given interval (the base outcome), 1 if she gives birth to a son in that interval, and 2 
if she gives birth to a daughter in that interval.

% The distribution of spacing is captured by the survival curve, which shows the probability 
% of not having had a birth yet by spell duration, for a given set of explanatory variables.
% The survival curve at time $t$ is 
% \begin{equation}
% \label{eq:survival}
% S_{t} 
% = 
% \prod_{d=1}^t
% \left(
% \frac{ 1 }
% {1 + \sum_{l=2}^2 \exp(D_j(t) + \alpha_{ld}'\mathbf{Z}_{kd} + \beta_l'\mathbf{X}_{k})}
% \right).
% \end{equation}
% 
% Because the probability of ever having a next birth varies across groups, a direct 
% comparison of standard survival curves tells us little about how the spread of sex 
% selection affects birth spacing across groups.
% I, therefore, condition on the predicted likelihood of parity progression when examining 
% birth spacing measures, such as the average duration to a birth.
% The reliability of this approach depends on whether the spell length covered is 
% sufficiently long that few women are likely to give birth after the spell cut-off.
% I discuss the choice of spell length below.%
% \footnote{
% It is important to note that the approach is not the same as merely calculating 
% the birth spacing measures for women who already have a given 
% parity child in the survey because that number does not take into account
% the censoring of spells that will eventually lead to a birth.
% }
% In addition, I present graphs of survival curves conditional on parity progression,
% which therefore begin at 100\% and end at 0\%.

I use the model to predict birth interval measures, the probability of having a birth, 
and the sex ratio.%
\footnote{
The interpretation of competing risk model coefficients is challenging \citep{thomas96}.
}
The sex ratio is based on the estimated percentage of children born at each $t$ that are 
boys, $\hat{Y}$:  
\begin{equation}
\hat{Y}_t 
= 
\frac{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )}
{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t) + P(g_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )} 
\times 100,
\label{eq:probability_son}
\end{equation}
where $b$ and $g$ are the predicted probabilities of having a boy and a girl, respectively,
conditional on the explanatory variables and not having had a child before $t$.
Combining the percentage of boys and the likelihood of exiting the spell with a birth
across all $t$ gives the predicted percent of births that are boys over the entire spell.%
\footnote{
With T=2, if 54 and 66 percent of the births are boys and the likelihoods of 
giving birth 20 and 40 percent, then $\frac{54*0.2+66*0.4}{0.2+0.4} = 62$ percent of the
births are boys.
}
The predicted sex ratio shown is the weighted average of the individual predicted sex 
ratios, using the parity progression probabilities as weights.

>>>>

Equation (\ref{eq:hazard}) is equivalent to the logistic hazard model and has the same 
likelihood function as the multinomial logit model \citep{allison82,jenkins95}.
Hence, splitting spells into smaller intervals---here equal to three months---and 
treating them as observations, I can estimate the model using a standard multinomial 
logit model. 
In the reorganized data, the outcome variable is 0 if the woman does not have a child 
in a given interval (the base outcome), 1 if she gives birth to a son in that interval, 
and 2 if she gives birth to a daughter in that interval.

I use the model to predict birth interval measures, the probability of having a birth, 
and the sex ratio.%
\footnote{
The interpretation of competing risk model coefficients is challenging \citep{thomas96}.
}
The predicted sex ratio is the weighted average of individual predicted sex ratios, using 
parity progression probabilities as weights.
For an individual woman, the estimated percent of births that are boys at $t$, $\hat{Y_t}$, is:  
\begin{equation}
\hat{Y}_t 
= 
\frac{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )}
{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t) + P(g_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )} 
\times 100,
\label{eq:probability_son}
\end{equation}
where $b$ and $g$ are the predicted probabilities of having a boy and a girl, respectively,
conditional on the explanatory variables and not having had a child before $t$.
Weighting the percentages boys with the likelihood of exiting the spell with a birth across all 
$t$ gives the predicted percent of boys over the entire spell for an individual.%
\footnote{
Imagine $T=2$. 
If 54 and 66 percent of births are boys and the likelihood of giving birth 20 and 40 percent, 
then the predicted sex ratio is $\frac{54*0.2+66*0.4}{0.2+0.4} = 62$ percent boys. 
}
